---
title: "Starbuck Customer Analysis"
author: "Trang Hoang"
date: "2025-02-22"
output: 
    html_document:
    toc: true
    toc_depth: '3'
    theme: darkly
    highlight: zenburn
    self_contained: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{css, echo=FALSE}
.ans {
   background-color:rgb(248,248,248); 
  padding: 1em; 
  border: .1em solid #CCC;
  border-radius:.2em;
  color:purple;
  margin: 1em 0;
}
```

```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(gridExtra)
library(viridis)  
library(ggthemes)
```

################################################# 

# 1. CUSTOMER DEMOGRAPHICS ANALYSIS

################################################# 

This section analyzes the age and membership distribution of customers. We'll create age groups and visualize their relationship with membership tiers.

```{r}
demographics <- read.csv("customer_demographics.csv")
stores <- read.csv("store_data.csv")
purchases <- read.csv("purchase_history.csv")
loyalty <- read.csv("loyalty_data.csv")
```

Create customer segments

```{r}
demographics <- demographics %>%
  mutate(income = as.numeric(income))
customer_segments <- demographics %>%
  left_join(loyalty, by = "customer_id") %>%
  mutate(
    age_group = cut(age, breaks = c(0, 25, 35, 50, 100),
                    labels = c("18-25", "26-35", "36-50", "50+")),
    income_group = cut(income, breaks = c(0, 30000, 60000, 90000, Inf),
                      labels = c("Low", "Medium", "High", "Very High"))
  )
```

Create visualization of age distribution

```{r}
ggplot(customer_segments, aes(x = age_group, fill = membership_tier)) +
  geom_bar(position = "dodge") +
  scale_fill_viridis(discrete = TRUE, option = "D") +
  labs(title = "Customer Age Distribution by Membership Tier",
       subtitle = "Understanding age group representation across membership levels",
       x = "Age Group", 
       y = "Number of Customers",
       fill = "Membership Tier") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

Assessment:

################################################# 

# 2. PURCHASE TIMING ANALYSIS

################################################# 

This section examines when purchases occur throughout the day. We'll analyze hourly patterns to identify peak times.

```{r}
purchase_patterns <- purchases %>%
  mutate(
    date = as.Date(date),
    month = floor_date(date, "month"),
    hour = hour(as.POSIXct(time, format = "%H:%M:%S"))
  )
```

Create hourly distribution plot

```{r}
ggplot(purchase_patterns, aes(x = hour)) +
  geom_histogram(aes(fill = ..count..), bins = 24) +
  scale_fill_viridis() +
  labs(title = "Hourly Purchase Distribution",
       subtitle = "Understanding peak hours of business",
       x = "Hour of Day (24-hour format)",
       y = "Number of Transactions") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(0, 23, 2))
```

Assessment:

################################################# 

# 3. PRODUCT PERFORMANCE ANALYSIS

################################################# 

This section analyzes which products generate the most revenue We'll create a ranking of drinks by total revenue

```{r}
product_mix <- purchases %>%
  group_by(drink_type) %>%
  summarize(
    total_revenue = sum(total_amount),
    count = n(),
    avg_price = mean(total_amount)
  ) %>%
  arrange(desc(total_revenue))
```

Create product revenue visualization

```{r}
ggplot(product_mix, aes(x = reorder(drink_type, total_revenue), y = total_revenue)) +
  geom_bar(stat = "identity", aes(fill = total_revenue)) +
  scale_fill_viridis() +
  coord_flip() +
  labs(title = "Revenue by Drink Type",
       subtitle = "Ranking of drinks by total revenue",
       x = "Drink Type",
       y = "Total Revenue ($)") +
  theme_minimal() +
  theme(legend.position = "none")
```

Assessment:

################################################# 

# 4. CUSTOMER VALUE ANALYSIS

################################################# 

This section identifies high-value customers and their characteristics We'll analyze the relationship between age, spending, and visit frequency

```{r}
customer_value <- purchase_patterns %>%
  group_by(customer_id) %>%
  summarize(
    total_spent = sum(total_amount),
    visit_count = n()
  ) %>%
  left_join(demographics, by = "customer_id") %>%
  mutate(
    value_segment = cut(total_spent, 
                       breaks = quantile(total_spent, probs = seq(0, 1, 0.25)),
                       labels = c("Low", "Medium", "High", "Very High"))
  )
```

Create customer value visualization

```{r}
ggplot(customer_value, aes(x = age, y = total_spent, color = value_segment)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linetype = "dashed") +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "Customer Value Distribution by Age",
       subtitle = "Understanding high-value customer characteristics",
       x = "Age",
       y = "Total Amount Spent ($)",
       color = "Value Segment") +
  theme_minimal()
```

Assessment:

################################################# 

# 5. STORE PERFORMANCE ANALYSIS

################################################# 

This section compares performance across different store locations We'll analyze revenue and transaction patterns by store type

```{r}
store_performance <- purchases %>%
  group_by(store_id) %>%
  summarize(
    total_revenue = sum(total_amount),
    avg_transaction = mean(total_amount),
    n_transactions = n()
  ) %>%
  left_join(stores, by = "store_id")
```

Create store performance visualization

```{r}
ggplot(store_performance, 
             aes(x = reorder(location_type, total_revenue), 
                 y = total_revenue, 
                 fill = store_type)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(title = "Revenue by Store Location",
       subtitle = "Comparing performance across store types",
       x = "Location Type",
       y = "Total Revenue ($)",
       fill = "Store Type") +
  theme_minimal()
```

Assessment:

################################################# 

# 6. SUMMARY STATISTICS

################################################# 

Generate overall business performance metrics

```{r}
summary_stats <- list(
  "Total Customers" = nrow(demographics),
  "Total Revenue" = sum(purchases$total_amount),
  "Average Transaction" = mean(purchases$total_amount),
  "Most Popular Drink" = names(which.max(table(purchases$drink_type))),
  "Best Performing Store" = store_performance$location_type[which.max(store_performance$total_revenue)]
)
summary_stats
```
